---
title: "Assignment 3"
output: html_document
date: "2026-03-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 1: Use sum() to determine the sum of numbers from 2000 to 20000. Provide your code and the actual sum in your answer.

```{r}
sum(2000:20000)
```

## Exercise 2: In one or two sentences, describe what the above code snippet did.

```{r}
d<-"I just want to be deleted"
d
ls()
#list the object "d"
rm(d)
#remove the object "d"
```

## Exercise 3: Add the contents of a and b together and place the results in a new object. Examine the result. Include your code. Try using both sum() and +; do you get different results? If so, why?

```{r}
a <- 5
b <- 2:20

c1 <- a + b
c1

c2 <- sum(a, b)
c2

#Using '+', R adds 'a' to each element of 'b', which gives a vector from 7 to 25.  
#Using 'sum()', R adds all values in 'a' and 'b' together, which gives 214.  
#The results are different because '+' performs element-wise addition, while 'sum()' returns one total.
```

## Exercise 4: What is the sum of the 5th through 10th element of object b? Provide your code and the sum.

```{r}
b <- 2:20
sum(b[5:10])
```

## Exercise 5: What is the sum of the 3rd, 8th, and 10th element of b?

```{r}
b <- 2:20
sum(b[c(3, 8, 10)])
```

## Exercise 6: When extracting from a 2-dimensional object, which number specifies rows and which specifies columns? What does m[3,] do? How can you extract the 3rd, 4th and 5th columns of m together as one object?

```{r}
m <- matrix(data = 1:25, ncol = 5, byrow = TRUE)
m[3,1]
m[, 3:5]

#In 'm[row, column]', the first number gives the row and the second number gives the column.  
#So 'm[3,1]' returns the value in the 3rd row and 1st column, which is 11.  
#To extract the 3rd, 4th, and 5th columns together, use 'm[, 3:5]'.
```

## Exercise 7: What does the cbind command do? How about rbind? Create a new object “n” where the first row is a new row of numbers (your choice) and the following rows are the rows from matrix m. Optional: do the same but reverse the order of the rows from matrix m.

```{r}
new_col <- cbind(m, 101:105)
new_col
#The code above adds a new column to the right side of matrix 'm'. The new column contains the values 101, 102, 103, 104, and 105.

n <- (50:54)
new <- rbind(n, m)
new
#The code above creates a new object 'new' by adding a new row to the top of matrix 'm'.
```

